---
title: "Project 1"
author: "Maria Garcia"
date: "2024-06-07"
output: 
  ioslides_presentation: 
    css: 'project1_scrolling_slides.css'

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(ggplot2)
library(corrplot)
library(reshape2)
library(tidyverse)
library(caret)
library(pROC)
library(cluster)
library(factoextra)
```

<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
color: black;
font-family: Modern Computer Roman;
}
slides > slide.title-slide hgroup h1 {
color: #8C1D40; <!--the maroon color-->
}
h2 {
color: #8C1D40; <!-- the maroon color-->
}
</style> <!-- end of defining font in various parts of slides -->

slides > slide {
  overflow-x: auto !important;
  overflow-y: auto !important;
}

## Introduction {.bigger}

- Analyzed customer data from Waze (crowdsourced navigation app).\n

- **Goal**: Identify factors predicting customer churn using visual and statistical methods.

## Importing the Data {.smaller}
```{r, echo=TRUE, warning=FALSE, quiet = TRUE, message=FALSE}
waze_data <- read_csv("/Users/mariagarcia/Desktop/DAT301/waze_app_dataset.csv")
head(waze_data)
```


## Preprocessing

- Renamed "label" to "customer_status".
- Removed ID column.
- Removed rows with missing customer_status.

```{r eval=TRUE, echo=TRUE}
colnames(waze_data)[colnames(waze_data) == "label"] <- "customer_status"
colnames(waze_data)

sapply(waze_data, function(x) sum(is.na(x)))
nrow(waze_data)

waze_data_cleaned <- waze_data[!is.na(waze_data$customer_status), ]
sapply(waze_data_cleaned, function(x) sum(is.na(x)))
nrow(waze_data_cleaned)


waze_data_cleaned <- waze_data_cleaned %>% select(-ID)
colnames(waze_data_cleaned)
```

## Distribution of Customer Status

- 82.3% retained, 17.7% churned.

```{r eval=TRUE, echo=TRUE}
churn_data <- waze_data_cleaned[!is.na(waze_data_cleaned$customer_status), ]
table(waze_data_cleaned$customer_status)
prop.table(table(waze_data_cleaned$customer_status))
```

## Correlation Analysis

Correlation heatmap of numerical features.

```{r, eval=TRUE, warning=FALSE, quiet = TRUE, message = FALSE, echo=TRUE}
churn_data$customer_status <- as.factor(churn_data$customer_status)
numeric_cols <- unlist(lapply(churn_data, is.numeric))
churn_numeric <- churn_data[, numeric_cols]
cor_matrix <- cor(churn_numeric)
corrplot(cor_matrix, method = "circle", type = "lower", tl.coll = "black", tl.srt = 45)
```

## Density Plots

Visualizing distribution of features by customer status.

```{r, eval = TRUE, echo=FALSE, warning=FALSE, quiet = TRUE, message = FALSE}
selected_features <- c("n_days_after_onboarding", "activity_days", "driving_days")

plot_list <- list()
for (feature in selected_features) {
  p <- ggplot(waze_data_cleaned, aes(x = !!sym(feature), color = customer_status, fill = customer_status)) +
    geom_density(alpha = 0.5) +
    labs(title = paste("Density Plot of", feature, "by Customer Status"), x = feature, y = "Density") +
    theme_minimal() +
    theme(plot.title = element_text(size = 10))
  plot_list[[feature]] <- p
  
}

plot_list

```

## Box Plots 

```{r, eval=TRUE, echo=FALSE, warning=FALSE, quiet = TRUE, message = FALSE}
numerical_features <- waze_data_cleaned %>% select_if(is.numeric) %>% names()
for (feature in numerical_features) {
  p <- ggplot(waze_data_cleaned, aes(x = customer_status, y = !!sym(feature), fill = customer_status)) +
    geom_boxplot() +
    labs(title = paste("Boxplot of", feature, "by Customer Status"),
         x = "Customer Status",
         y = feature) +
    theme_minimal()
  print(p)
}
```


## Logistic Regression Model

```{r, eval=TRUE, quiet=TRUE, message = FALSE, warning=FALSE, echo = TRUE}
waze_data_cleaned$customer_status <- as.factor(waze_data_cleaned$customer_status)
model <- glm(customer_status ~ sessions + drives + total_sessions + 
             n_days_after_onboarding + total_navigations_fav1 + 
             total_navigations_fav2 + driven_km_drives + 
             duration_minutes_drives + activity_days + 
             driving_days + device, 
             data = waze_data_cleaned, family = binomial)

summary(model)
```

## Summary of Model Results 
1. n_days_after_onboarding: Longer onboarding duration increases the likelihood of churn.
2. total_navigations_fav1: More navigations to the first favorite location decrease churn likelihood.
3. total_navigations_fav2: More navigations to the second favorite location decrease churn likelihood.
4. duration_minutes_drives: Longer drive durations reduce churn likelihood.
5. activity_days: More activity days increase retention likelihood.
6. driving_days: More driving days reduce churn likelihood.

## Conclusion 

- Increased engagement reduces churn.
- Longer onboarding duration increases churn.
- Maintain customer engagement to improve retention.
